---
title: "Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
devtools::load_all()
library(tidyverse)
library(patchwork)
```

### Construct a GP emulator for Energy Systems Model.

We proceed to construct a GP emulator with a linear form of the regression function, $h(\boldsymbol{x})=(1, x_1, x_2, x_3, x_4, x_5)$, squared exponential correlation function (kernel) and nugget term $\tau^2=1$.

```{r, warning=FALSE, results="hide"}
n_design <- dim(w_design)[1]
design <- as.matrix(cbind(w_design, z_design))
m.E <- RobustGaSP::rgasp(design = design,
             trend = cbind(rep(1, n_design), design),
             response = y_design, nugget = 0.1,
             kernel_type = "pow_exp",
             alpha = rep(2, 5))
```

We use RobustGaSP package to fit a GP emulator. We proceed to perform validations.

```{r}

n_valid <- dim(w_valid)[1]
valid <- as.matrix(cbind(w_valid, z_valid))
LOO.m.E <- RobustGaSP::leave_one_out_rgasp(m.E)

predict.m.E <- RobustGaSP::predict(m.E, valid,
                       testing_trend = cbind(rep(1, n_valid), valid))
```

```{r, fig.cap = "Traditional Leave One Out (top row) and cross-validation (bottom row) plots for energy systens model emulator."}
loo_energy <- tibble(x = LOO.m.E$mean,
                     y = y_design,
                     y_lower = LOO.m.E$mean-2*LOO.m.E$sd,
                     y_upper = LOO.m.E$mean+2*LOO.m.E$sd)%>%
  mutate(perform = if_else(y_lower <= y & y <= y_upper, 1, 0)) %>%
  mutate(perform = as.factor(perform)) %>%
  ggplot() +
  geom_errorbar(aes(x = x, ymin = y_lower, ymax = y_upper),
                width = 20) +
  geom_point(aes(x = x, y = y, colour = perform)) +
  scale_color_manual(" ", breaks = c(0, 1),
                     values = c(wesanderson::wes_palettes$FantasticFox1[5],                 wesanderson::wes_palettes$FantasticFox1[3]),
                     labels = c("out", "in")) +
  labs(x = "Prediction", y = "True values")


valid_energy <- tibble(x = predict.m.E$mean,
                       y = y_valid,
                       y_lower = predict.m.E$mean-2*predict.m.E$sd,
                       y_upper = predict.m.E$mean+2*predict.m.E$sd)%>%
  mutate(perform = if_else(y_lower <= y & y <= y_upper, 1, 0)) %>%
  mutate(perform = as.factor(perform)) %>%
  ggplot() +
  geom_errorbar(aes(x = x, ymin = y_lower, ymax = y_upper),
                width = 20) +
  geom_point(aes(x = x, y = y, colour = perform)) +
  scale_color_manual(" ",
                     breaks = c(0, 1),
                     values = c(wesanderson::wes_palettes$FantasticFox1[5],                                wesanderson::wes_palettes$FantasticFox1[3]),
                     labels = c("out", "in")) +
  labs(x = "Prediction", y = "True values")


loo_energy/valid_energy

```

### Produce components necessary for the computation of linked emulator.

```{r}
d <- 3 # number of random inputs
d_t <- 5 # total number of inputs

params <- list(delta_par = 1/m.E@beta_hat,
               sigma = sqrt(m.E@sigma2_hat),
               beta = c(m.E@theta_hat[2:(d+1)],
                        m.E@theta_hat[-(2:(d+1))]),
               nugget = m.E@nugget)

m = dim(m.E@input)[1]

A <- CovMatrix(m.E@input, params$delta_par) +
    diag(params$nugget, ncol = m, nrow = m)
QA <- chol(A)
R <- solve(A)

diff <-  m.E@output- as.matrix(m.E@input[, 1:d])%*%as.matrix(c(params$beta[1:d]),
                                                              nrow = d) -
    cbind(rep(1, m), as.matrix(m.E@input[, -(1:d)]))%*%as.matrix(c(params$beta[-(1:d)]),
                                                              nrow = (d_t+1)-d)

Ldiff <- backsolve(QA, diff, transpose=TRUE)
FastParts<- list(A = A, QA = QA, Ldiff = Ldiff,R = R, diff = diff)
```

### Predictions produced by linked emulator.

We proceed to compute mean and variance of the linked emulator using two functions `mean_fun` and `var_fun`.

```{r}
n2 <- dim(mean_pred)[1] # specify the size of the validation data set.

# compute mean values
linked_mean_vec <- sapply(1:n2, function(x)
  mean_fun(w = as.matrix(w_design),
           z = c(z_valid$z1[x], z_valid$z2[x]),
           z_design = as.matrix(z_design),
           y =  y_design,
           mu = c(mean_pred$mu1[x],mean_pred$mu2[x],mean_pred$mu3[x]),
           Sigma = matrix(c(covar_pred$var1[x], 0, 0, 
                            0, covar_pred$var2[x], covar_pred$covar23[x], 
                            0, covar_pred$covar23[x], covar_pred$var3[x]), 
                          nrow = 3, ncol = 3), 
           params = params, FastParts = FastParts, ind = "FALSE"))

# compute variance values.
linked_var_vec <- sapply(1:n2, function(x)
  var_fun(w = as.matrix(w_design),
           z = c(z_valid$z1[x], z_valid$z2[x]),
           z_design = as.matrix(z_design),
           y =  y_design,
           mu = c(mean_pred$mu1[x],mean_pred$mu2[x],mean_pred$mu3[x]),
           Sigma = matrix(c(covar_pred$var1[x], 0, 0, 
                            0, covar_pred$var2[x], covar_pred$covar23[x], 
                            0, covar_pred$covar23[x], covar_pred$var3[x]), 
                          nrow = 3, ncol = 3), 
           params = params, FastParts = FastParts, ind = "FALSE"))


```

```{r, fig.cap = "One-step ahead predictions: black points correspond to the total operational costs produced by the energy system simulator. The red line and the red dotted lines correspond to the forecasts produced by the linked model.", fig.width = 10,  fig.heigh = 5}
valid.energy.linked <- tibble(date = 1:n2,
                              y_valid = y_valid,
                              pred_m = linked_mean_vec,
                              pred_l = linked_mean_vec- 2*sqrt(linked_var_vec),
                              pred_u = linked_mean_vec + 2*sqrt(linked_var_vec))


pred_plot_linked <- ggplot(valid.energy.linked[-1, ]) +
  geom_line(aes(x = date, y = pred_l),
            color = "red", linetype = "dashed") +
  geom_line(aes(x = date, y = pred_u),
            color = "red", linetype = "dashed") +
  geom_line(aes(x = date, y = pred_m), color = "red") +
  geom_point(aes(x = date, y = y_valid)) +
  ylab("Total operational costs (£)") +
  xlab("") 

pred_plot_linked
```





